---
output: 
  html_document: 
    fig_height: 6
    fig_width: 12
---
# **EDA of YouTube Trending Statistics**

#### **Exploratory Data Analysis** | Udacity **Data Analyst Nanodegree**

<br>
<h3 style="text-align:left;"><b>Karra Anand</b></h3>
<h3 style="text-align:left;">26<sup>th</sup> March 2018</h3>
<h4 style="text-align:left;">Version: 1.1</h3>
___
___


<h2>Table of Contents</h2>
<ul>
<li><a href="#about-the-dataset"><b>About the dataset</b></a></li>
<li><a href="#preprocessing"><b>Preprocessing</b></a></li>
<a href="#cleaning-irrelevant-columns">&nbsp;&nbsp;&nbsp;&nbsp;
Cleaning irrelevant columns
</a><br>
<a href="#converting-into-a-data-table">&nbsp;&nbsp;&nbsp;&nbsp;
Converting into a data.table
</a><br>
<a href="#adding-category-names">&nbsp;&nbsp;&nbsp;&nbsp;
Adding category names
</a><br>
<a href="#adding-number-of-videos-trending">&nbsp;&nbsp;&nbsp;&nbsp;
Adding number_of_videos_trending
</a><br>
<a href="#adding-and-sorting-by-days-on-trending">&nbsp;&nbsp;&nbsp;&nbsp;
Adding and sorting by days_on_trending
</a><br>
<a href="#creating-a-data-table-sorted-by-number-of-videos-trending">
&nbsp;&nbsp;&nbsp;&nbsp;
Creating a data.table sorted by number_of_videos_trending
</a><br>
<a href="#removing-duplicate-entries">&nbsp;&nbsp;&nbsp;&nbsp;
Removing duplicate entries
</a>
<li><a href="#univariate-plots-section"><b>Univariate Plots section</b></a></li>
<a href="#univariate-top100-trending-videos">&nbsp;&nbsp;&nbsp;&nbsp;
Top 100 Trending Videos
</a>
<li><a href="#bivariate-plots-section"><b>Bivariate Plots Section</b></a></li>
<a href="#bivariate-top100-trending-videos">&nbsp;&nbsp;&nbsp;&nbsp;
Top 100 Trending Videos
</a>
<li><a href="#bivariate-analysis"><b>Bivariate Analysis</b></a></li>
<li><a href="#multivariate-plots-section"><b>Multivariate Plots Section
</b></a></li>
<li><a href="#multivariate-analysis"><b>Multivariate Analysis</b></a></li>
<li><a href="#final-plots-and-summary"><b>Final Plots and Summary</b></a></li>
<a href="#plot-one">&nbsp;&nbsp;&nbsp;&nbsp;
Plot One
</a><br>
<a href="#description-one">&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;
Description One
</a><br>
<a href="#plot-two">&nbsp;&nbsp;&nbsp;&nbsp;
Plot Two
</a><br>
<a href="#description-two">&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;
Description Two
</a><br>
<a href="#plot-three">&nbsp;&nbsp;&nbsp;&nbsp;
Plot Three
</a><br>
<a href="#description-three">&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;
Description Three
</a>
<li><a href="#reflection"><b>Reflection</b></a></li>
</ul>

___
___

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(data.table)
library(corrplot)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
yt_trending = read.csv('USvideos.csv')
```


<a id='about-the-dataset'></a>

## About the dataset

YouTube (the world-famous video sharing website) maintains a list of the top 
trending videos on the platform. This dataset is a daily record of the top 
trending YouTube videos. <br>
This dataset includes several months (and counting) of data on daily trending 
YouTube videos. Data is included for the USA, with up to 200 listed trending 
videos per day. <br>
More details about this dataset are present in `About_dataset.txt` included with 
this project

___
___

<a id='preprocessing'></a>

## Preprocessing
Before we move on to plotting and analyzing the data, let us see if the data 
requires any cleaning.
<br>
<br>
The first 2 rows of each of the columns of the dataset are as follow:
```{r echo=FALSE,preproceessing}
head(yt_trending,n=2)
```
<br>

> In many of the following data cleaning steps only the code but not the output 
> is printed to prevent repeated printing of the same dataset with minor 
> modifications. The final dataset obtained after the data cleaning is printed 
> at the end of this section.

___

<a id='cleaning-irrelevant-columns'></a>

### Cleaning irrelevant columns
As we are only interested in exploratory analysis of the data, we remove the 
columns of **tags**, **thumbnail_link** and **description** since they are 
irrelevant to us.
```{r echo=FALSE, preprocessing_2}

yt_trending <- select(yt_trending,-tags,-thumbnail_link,
                      -description)

head(yt_trending)

dim(yt_trending)
```

___

<a id='converting-into-a-data-table'></a>

### Converting into a data.table
From the dimensions of the dataframe in the above output, we have
the data for 23,362 videos(assuming there are no duplicates across 13 features; 
we shall investigate  this the following sections). Hence, we use the data.table
 data structure to store our data as it superior to a dataframe for add, remove,
 update, join etc. operations.
```{r echo=TRUE, preprocessing_3}
yt_trending <- data.table(yt_trending)
```

___

<a id='adding-category-names'></a>

### Adding category names
Further we observe that, we have names of each of the **category_id** (as part
of `US_category_id.json` file). Hence, adding the column for **category_name**.

```{r echo=TRUE, preprocessing_4}
category_id <- c(1,2,10,15,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,
                  34,35,36,37,38,39,40,41,42,43,44)
category_name <- c("Film & Animation","Autos & Vehicles","Music",
                    "Pets & Animals","Sports","Short Movies",
                    "Travel & Events","Gaming","Videoblogging",
                    "People & Blogs","Comedy","Entertainment",
                    "News & Politics","Howto & Style","Education",
                    "Science & Technology",
                    "Nonprofits & Activism","Movies","Anime/Animation",
                    "Action/Adventure","Classics","Comedy",
                    "Documentary","Drama","Family","Foreign","Horror",
                    "Sci-Fi/Fantasy","Thriller","Shorts","Shows","Trailers")

category_id_names <- data.frame(category_id,category_name)

yt_trending <- merge(yt_trending,category_id_names)

yt_trending <- yt_trending[order(yt_trending$views,decreasing = TRUE),]
```

___

<a id='adding-number-of-videos-trending'></a>

### Adding number_of_videos_trending
The number of videos each channel has in the trending videos is added to the 
dataset.

```{r echo=TRUE, preprocessing_5}
number_of_videos_trending_channel_title <- 
  (count(yt_trending,yt_trending$channel_title))[1]
number_of_videos_trending_count <-
  (count(yt_trending,yt_trending$channel_title))[2]
number_of_videos_trending <- data.table(number_of_videos_trending_channel_title,
                                        number_of_videos_trending_count)

setnames(number_of_videos_trending,old="yt_trending$channel_title",
         new="channel_title")
setnames(number_of_videos_trending,old="n",new="number_of_videos_trending")

yt_trending <- merge(yt_trending,number_of_videos_trending)
```

___

<a id='adding-and-sorting-by-days-on-trending'></a>

### Adding and sorting by days_on_trending
We are most interested here in analyzing the trends in the trending videos,
and we do not have a columnn to indicate the number of days a video was on 
trending, we add that column here.

```{r echo=TRUE, preprocessing_6}
days_on_trending_video_id <- (count(yt_trending,yt_trending$video_id))[1]
days_on_trending_num_of_days <- (count(yt_trending,yt_trending$video_id))[2]
days_on_trending <- data.table(days_on_trending_video_id,
                               days_on_trending_num_of_days)

setnames(days_on_trending,old="yt_trending$video_id",new="video_id")
setnames(days_on_trending,old="n",new="days_on_trending")

yt_trending <- merge(yt_trending,days_on_trending,by="video_id")

yt_trending <- yt_trending[order(yt_trending$days_on_trending,
                                 yt_trending$views,
                                 decreasing = TRUE),]
```

___

<a id='creating-a-data-table-sorted-by-number-of-videos-trending'></a>

### Creating a data.table sorted by number_of_videos_trending
As per the requirement in the Bivariate Plots section for analysing the top 
trending channels.

```{r echo=TRUE, preprocessing_7}
yt_trending_by_number_of_videos_trending <-
  yt_trending[order(yt_trending$number_of_videos_trending,
                    yt_trending$views,
                    decreasing = TRUE),]
```
___

<a id='removing-duplicate-entries'></a>

### Removing duplicate entries
We sort the videos in decreasing order of their number of days on trending.

```{r echo=TRUE, preprocessing_8}
yt_trending <- unique(yt_trending, by = c("video_id"))
yt_trending_by_number_of_videos_trending <- unique(
  yt_trending_by_number_of_videos_trending,
  by=c("channel_title"))
```

___
___

<a id='univariate-plots-section'></a>

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots_stats}
dim(yt_trending)
names(yt_trending)
summary(yt_trending)

likes_colour <- "chartreuse2"
dislikes_colour <- "red"
comment_count_colour <- "darkorchid"
days_on_trending_colour <- "deeppink"
views_colour <- "dodgerblue4"
```
As we can see from the above output, we have the data of 4,712 videos and their 
15 features.


```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots}
category_names_plot <- ggplot(yt_trending,aes(x=category_name))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Category")

category_names_plot

summary(yt_trending$category_name)
```
Plotting the number of videos as per their category name, we can see a large 
amount of variation in the number of videos in each category. We can clearly 
observe that **Entertainment** and **Music** are the top 2 categories with 1141 
and 585 videos respectively closely followed by **Howto & Style** with 436 
videos. We can also see their respective shares out of the total.
<br>
<br>
Also, some categories like **Action/Adventure**, **Anime/Animation** etc. have 
no videos.

___

```{r echo=FALSE, fig_width = 12, fig_height = 12, Univariate_Plots2, warning=FALSE}
likes_plot <- ggplot(yt_trending,aes(x=yt_trending$likes))+
  geom_histogram(bins=50,fill="chartreuse2")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of likes)")

dislikes_plot <- ggplot(yt_trending,aes(x=yt_trending$dislikes))+
  geom_histogram(bins=50,fill="red")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of dislikes)")

views_plot <- ggplot(yt_trending,aes(x=yt_trending$views))+
  geom_histogram(bins=50,fill="dodgerblue4")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of views)")

comments_plot <- ggplot(yt_trending,aes(x=yt_trending$comment_count))+
  geom_histogram(bins=50,fill="darkorchid")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of comments)")

grid.arrange(views_plot,likes_plot,dislikes_plot,comments_plot,ncol=1)

print("Summary of views feature:")
summary(yt_trending$views)

print("Summary of likes feature:")
summary(yt_trending$likes)

print("Summary of dislikes feature:")
summary(yt_trending$dislikes)

print("Summary of comments feature:")
summary(yt_trending$comment_count)
```
This plot shows the log10 of the number of **views**, **likes**, **dislikes** and 
**comments** on the videos.
<br>
<br>
We can observe a distribution similar to that of a normal distribution. Further,
we can see that, the mean number of views is more that that of any other 
attribute.
<br>
<br>
The **mean**, **median** and **maximum** number of view are 1,278,000, 331,600 
and 149,400,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of likes are 39,720, 7,726 and 
3,094,000 repectively.
<br>
<br>
The **mean**, **median** and **maximum** number of dislikes are 2,598, 302 and 
1,674,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of comments are 4,976, 888 and 
1,362,000 respectively.

___

```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots3}
comments_disabled_overall_plot <- ggplot(yt_trending,
                                         aes(x=yt_trending$comments_disabled))+
  geom_bar()+
  xlab("Comments disabled")

ratings_disabled_overall_plot <- ggplot(yt_trending,
                                        aes(x=yt_trending$ratings_disabled))+
  geom_bar()+
  xlab("Ratings disabled")

grid.arrange(comments_disabled_overall_plot,ratings_disabled_overall_plot,
             ncol=2)

print("Summary of videos with comments disabled")
summary(yt_trending$comments_disabled)

print("Summary of videos with ratings disabled")
summary(yt_trending$ratings_disabled)
```

So we have only 79 videos which have their comments disabled and only 26 videos 
with their ratings disabled.

___

```{r echo=FALSE, warning=FALSE, Univariate_Plots4}
ggplot(yt_trending,aes(x=yt_trending$days_on_trending))+
  geom_bar(fill="deeppink")+
  xlab("Days on trending")

summary(yt_trending$days_on_trending)
```
We can observe a bimodal type of distribution for the **days on trending**.
<br>
<br>
With videos trending for an average of about 5 days and a maximum of 14 days.

___

<a id='univariate-top100-trending-videos'></a>

### Top 100 Trending Videos

```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots5,warning=FALSE}
top100_trending <- yt_trending[1:100]

top100_category_names_plot <- ggplot(top100_trending,aes(x=category_name))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Categories in Top 100 trending videos")

grid.arrange(top100_category_names_plot,category_names_plot)
```
Videos of some categories like **Auto & Vehicles**, **Comedy** etc. are not even
present in the top 100 trending videos.

<br>
```{r echo=FALSE, warning=FALSE, Univariate_Plots6}
top100_likes_plot <- ggplot(top100_trending,aes(x=top100_trending$likes))+
  geom_histogram(bins=50,fill="chartreuse2")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10 of number of likes on videos")

top100_dislikes_plot <- ggplot(top100_trending,aes(x=top100_trending$dislikes))+
  geom_histogram(bins=50,fill="red")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of dislikes)")

top100_views_plot <- ggplot(top100_trending,aes(x=top100_trending$views))+
  geom_histogram(bins=50,fill="dodgerblue4")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of views)")

top100_comments_plot <- ggplot(top100_trending,
                               aes(x=top100_trending$comment_count))+
  geom_histogram(bins=50,fill="darkorchid")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of comments)")


grid.arrange(top100_views_plot,
             top100_likes_plot,
             top100_dislikes_plot,
             top100_comments_plot,
             ncol=1)

print("Summary of views feature:")
summary(top100_trending$views)

print("Summary of likes feature:")
summary(top100_trending$likes)

print("Summary of dislikes feature:")
summary(top100_trending$dislikes)

print("Summary of comments feature:")
summary(top100_trending$comment_count)
```

For the top 100 trending videos we cannot see a definitive distribution as we 
could for these same features for the entire distribution.
<br>
<br>
The **mean**, **median** and **maximum** number of view are 3,411,000, 1,665,000 
and 45,940,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of likes are 105,600, 49,440 and 
822,000 repectively.
<br>
<br>
The **mean**, **median** and **maximum** number of dislikes are 5,052, 1,600 and 
165,100 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of comments are 11,980, 4,667 
and 203,900 respectively.
<br>
<br>
Comparing these statistics to those for the entire distribution, **all of these 
statistics are higher than those for the entire distribution**.

___

```{r echo=FALSE, warning=FALSE, Univariate_Plots7}
ggplot(top100_trending,aes(x=top100_trending$days_on_trending))+
  geom_bar(fill="deeppink")+
  xlab("Days on trending")

summary(top100_trending$days_on_trending)
```

Videos in the Top 100 trending videos trend for an average of 13 days with a 
maximum of 14 days.
___
___

<a id='univariate-analysis'></a>

## Univariate Analysis

### What is the structure of your dataset?
The dataset after preprocessing to remove duplicates, contains the data for 
4712 videos and 15 features (video_id, category_id, trending_date, title, 
channel_title, publish_time, views, likes, dislikes, comment_count, 
comments_disabled, ratings_disabled, video_error_or_removed, category_name, 
days_on_trending)
<br>
<br>
**Unordered factors: **category_id, category_name.
<br>
  **category_id: **1,2,10,15,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,
                  34,35,36,37,38,39,40,41,42,43,44
  <br>
  **category_name: **"Film & Animation","Autos & Vehicles","Music",
                    "Pets & Animals","Sports","Short Movies",
                    "Travel & Events","Gaming","Videoblogging",
                    "People & Blogs","Comedy","Entertainment",
                    "News & Politics","Howto & Style","Education",
                    "Science & Technology",
                    "Nonprofits & Activism","Movies","Anime/Animation",
                    "Action/Adventure","Classics","Comedy",
                    "Documentary","Drama","Family","Foreign","Horror",
                    "Sci-Fi/Fantasy","Thriller","Shorts","Shows","Trailers"
<br>
<br>
**Other observations:**

* **Entertainment** is the category with the highest number of video in 
trending with a total of 1141 videos.
* Average number of views on a trending video are 1,278,000.
* The maximum number of days for which a video stayed on trending is 14.
* Average number of views on a video in the top 100 trending videos is 
3,411,000.
* Most of the videos in the top 100 trending videos stay on trending for 13 
days.

___


### What is/are the main feature(s) of interest in your dataset?

The **number of views (views)**, **number of days the video is trending 
(days_on_trending)** and **category of the video (category_name)** are the main 
features of interest here.

___

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Although subject to the viewers' biases, number of likes, dislikes and comments 
can also help in understanding the video's position on the list of trending 
videos.

___

### Did you create any new variables from existing variables in the dataset?

Yes, the variable **category_name** was created to better interpret the 
category_id feature which has a direct correspondance with category_name. <br>
Also, the **days_on_trending** variable was created to keep track of the number 
of days the video was on trending.

___

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

One of the unusual observations were that there were no videos from some of the 
categories like **Action/Adventure**, **Anime/Animation** etc.

Some features like **tags**, **thubnail_link** and **description** were removed 
from the dataset as they took a lot of space on printing and were irrelevant to 
our analysis.

Log10 transform was applied in multiple plots to convey the scale and variation 
in the data as required.

Also, the same video which as trending on nultiple days was reduced to a single 
entry on the day it had the highest views.

___
___

<a id='bivariate-plots-section'></a>

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
corrplot_overall <- corrplot.mixed(corr = cor(yt_trending[,c("views",
                                         "likes",
                                         "dislikes",
                                         "comment_count",
                                         "days_on_trending"),with=F]))

ggpairs(yt_trending[,c("views",
"likes",
"dislikes",
"comment_count",
"days_on_trending")])

```

The above correlation matrix helps in identifying some of the interesting trends
 in the data.
 
We have a high correlation between

* the **number of views** and **number of likes**, and 
* the **number of dislikes** and **comment count**.

<br>
But, before we plot scatter plots to visualize these correlations, we have to 
normalize the data ranges of the above mentioned four features.

<br>
After normalizing to be in the range of [0,1]. We get the following output:


```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots2}
range01 <- function(x){(x-min(x))/(max(x)-min(x))}

normalized_values <- select(yt_trending,views,likes,dislikes,comment_count)

normalized_values$views <- range01(normalized_values$views)
normalized_values$likes <- range01(normalized_values$likes)
normalized_values$dislikes <- range01(normalized_values$dislikes)
normalized_values$comment_count <- range01(normalized_values$comment_count)
head(normalized_values)

views_likes_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$views,
                               y=normalized_values$likes))+
  geom_point(alpha=0.2,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlab("Normalized views")+
  ylab("Normalized likes")

dislikes_commentCount_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$dislikes,
                               y=normalized_values$comment_count))+
  geom_point(alpha=0.2,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  xlab("Normalized dislikes")+
  ylab("Normalized comment count")

grid.arrange(views_likes_plot,dislikes_commentCount_plot,ncol=2)
```

Using suitable limits for the X and Y axis:

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots3}
views_likes_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$views,
                               y=normalized_values$likes))+
  geom_point(alpha=0.2,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized likes")

dislikes_commentCount_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$dislikes,
                               y=normalized_values$comment_count))+
  geom_point(alpha=0.2,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.15)+
  ylim(0,0.15)+
  xlab("Normalized dislikes")+
  ylab("Normalized comment count")

grid.arrange(views_likes_plot,dislikes_commentCount_plot,ncol=2)
```

We can clearly observe the correlation that we found out previously using 
the correlation matrix.


We can see the variation in the features of **likes**, **dislikes**, **comment 
count** and **days on trending** in the following plots.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, Bivariate_Plots4}
likes_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$likes,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="chartreuse2")+
  geom_smooth()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(likes)")

dislikes_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$dislikes,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="red")+
  geom_smooth()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(dislikes)")

comments_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$comment_count,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="darkorchid")+
  geom_smooth()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(comment count)")

daysOnTrending_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$days_on_trending,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="deeppink")+
  geom_smooth()+
  scale_x_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  xlab("log10(views)")+
  ylab("days on trending")

grid.arrange(likes_views_plot,
             dislikes_views_plot,
             comments_views_plot,
             daysOnTrending_views_plot,
             ncol=2)
```

The notable but not considerable correlations like in between **views** and 
**dislikes** and that between **views** and **comment count** are visible here.

As the number of views increases, the other features also increase which is in 
agreement with our calculated statistics in the univariate plots section.

Moreover, only a small number of videos trend for 10 days or more.
<br>

Before we compare how the different features vary across the different 
categories, we take a look at the top 25 trending channels.
```{r echo=FALSE, Bivariate_Plots4_2}
ggplot(yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title,y=number_of_videos_trending))+
  geom_density()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_continuous(breaks=seq(0,140,10))+
  xlab("Channel title")+
  ylab("Number of videos trending")
```

**ESPN**, **Vox** and **Netflix** are the top 3 channels (in that order) in 
terms of number of trending videos.

How these trending channels vary across the categories is something that we will
take a look at in the multivariate plots section.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=15, Bivariate_Plots5}
category_views_boxplot <- ggplot(yt_trending,aes(x=category_name,y=views))+
  geom_boxplot(alpha=0.1,colour="dodgerblue4")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(views)")+
  xlab("category name")

category_likes_boxplot <- ggplot(yt_trending,aes(x=category_name,y=likes))+
  geom_boxplot(alpha=0.1,colour="chartreuse2")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(likes)")+
  xlab("category name")

category_dislikes_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                    y=dislikes))+
  geom_boxplot(alpha=0.1,colour="red")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(dislikes)")+
  xlab("category name")

category_commentCount_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                        y=comment_count))+
  geom_boxplot(alpha=0.1,colour="darkorchid")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(comment count)")+
  xlab("category name")

category_daysOnTrending_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                          y=days_on_trending))+
  geom_boxplot(alpha=0.1,colour="deeppink")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  ylab("days on trending")+
  xlab("category name")

category_numberOfVideosTrending <- ggplot(yt_trending,
                                           aes(x=category_name,
                                               y=number_of_videos_trending))+
  geom_boxplot(alpha=0.1)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,140,20))+
  ylab("Number of videos trending")+
  xlab("category name")

grid.arrange(category_views_boxplot,
             category_likes_boxplot,
             category_dislikes_boxplot,
             category_commentCount_boxplot,
             category_daysOnTrending_boxplot,
             category_numberOfVideosTrending,
             ncol = 2)

print("Summary for views feature:")
summary(yt_trending$views)

print("Summary for likes feature:")
summary(yt_trending$likes)

print("Summary for dislikes feature:")
summary(yt_trending$dislikes)

print("Summary for comment count feature:")
summary(yt_trending$comment_count)

print("Summary for days on trending feature:")
summary(yt_trending$days_on_trending)

print("Summary for number of trending videos feature:")
summary(yt_trending$number_of_videos_trending)
```

The categories of **Entertainment** and **Music** have very high values for all 
the features across the board.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

Overall, it seems that **Shows** is one category that consistently appears on
trending.

___

<a id='bivariate-top100-trending-videos'></a>

### Top 100 Trending Videos

In the univariated plots and analysis sections, we observed that the trends in 
the entire dataset were more pronounced in the Top 100 Trending Videos.

We check if this is true for the correlations that we observed at the beginning 
of the bivariate plots section.

```{r echo=FALSE, Bivariate_Plots6}
corrplot_top100 <- corrplot.mixed(corr = cor(top100_trending[,c("views",
                                         "likes",
                                         "dislikes",
                                         "comment_count",
                                         "days_on_trending"),with=F]))

ggpairs(top100_trending[,c("views",
                       "likes",
                       "dislikes",
                       "comment_count",
                       "days_on_trending")])
```


The changes in these correlations from those of the entire dataset are:

* Minor increase from 0.81 to 0.85 between **views** and **likes** 
* Drastic reduction from 0.83 to 0.36 between **dislikes** and **comment count**
* Of the other correlations, some has increased and some have decreased

Again before plotting, we first normalize the values of the features to be in 
the range [0,1].

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots7}
normalized_top100_trending <- select(top100_trending,
                                     views,
                                     likes,
                                     dislikes,
                                     comment_count)

normalized_top100_trending$views <- range01(top100_trending$views)
normalized_top100_trending$likes <- range01(top100_trending$likes)
normalized_top100_trending$dislikes <- range01(top100_trending$dislikes)
normalized_top100_trending$comment_count <- 
  range01(top100_trending$comment_count)

head(normalized_top100_trending)


top100_views_likes_plot <- ggplot(normalized_top100_trending,
                           aes(x=normalized_top100_trending$views,
                               y=normalized_top100_trending$likes))+
  geom_point(alpha=0.5,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized likes")

top100_views_dislikes_plot <- ggplot(normalized_top100_trending,
                           aes(x=normalized_top100_trending$views,
                               y=normalized_top100_trending$dislikes))+
  geom_point(alpha=0.5,colour="red")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized dislikes")

grid.arrange(top100_views_likes_plot,top100_views_dislikes_plot,ncol=2)
```

We can observe the correlation that we found from the correlation matrix above.

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=8, Bivariate_Plots8}
top100_likes_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$likes,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(likes)")

top100_dislikes_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$dislikes,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="red")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(dislikes)")

top100_comments_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$comment_count,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(comment count)")

top100_daysOnTrending_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$days_on_trending,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="deeppink")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  xlab("log10(views)")+
  ylab("days on trending")

grid.arrange(top100_likes_views_plot,
             top100_dislikes_views_plot,
             top100_comments_views_plot,
             top100_daysOnTrending_views_plot,
             ncol=2)
```

The data points are further away from the y=x line and hence overall, with 
**views** the other features are less correlated.

Also, videos in the Top 100 Trending Videos trend for either 12, 13 or 14 days 
only.

In this part, we try to find how the various features vary in the Top 100 
Trending Videos vary from those in the entire dataset.

The plots from **Top 100 Trending Videos are on the right** and those from the 
**entire dataset are on the left**.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=25, Bivariate_Plots9}
top100_category_views_boxplot <- ggplot(top100_trending,
                                        aes(x=category_name,y=views))+
  geom_boxplot(colour="dodgerblue4")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(views)")+
  xlab("category name")

top100_category_likes_boxplot <- ggplot(top100_trending,
                                        aes(x=category_name,y=likes))+
  geom_boxplot(colour="chartreuse2")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(likes)")+
  xlab("category name")

top100_category_dislikes_boxplot <- ggplot(top100_trending,aes(x=category_name,
                                                    y=dislikes))+
  geom_boxplot(colour="red")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(dislikes)")+
  xlab("category name")

top100_category_commentCount_boxplot <- ggplot(top100_trending,
                                               aes(x=category_name,
                                                        y=comment_count))+
  geom_boxplot(colour="darkorchid")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(comment count)")+
  xlab("category name")

top100_category_daysOnTrending_boxplot <- ggplot(top100_trending,
                                                 aes(x=category_name,
                                                          y=days_on_trending))+
  geom_boxplot(colour="deeppink")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  ylab("days on trending")+
  xlab("category name")

top100_category_numberOfTrendingVideos_boxplot <- ggplot(top100_trending,
                                                 aes(x=category_name,
                                                     y=number_of_videos_trending))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,140,20))+
  ylab("Number of videos trending")+
  xlab("category name")
```

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_0}
grid.arrange(category_views_boxplot,
             top100_category_views_boxplot,
             ncol=2)

summary(top100_trending$views)
```
The most notable category in this comparision is **Sports** which has a low 
median value in the entire dataset but the highest median, 1<sup>st</sup> and 
3<sup>rd</sup> quantile values in the Top 100 Trending Videos.

Surprisingly, the **Shows** category has completely disappeared in the Top 100 
Trending Videos set.

**Music** is the category with the second highest median value in the Top 100 
Trending Videos closely followed by **Film & Animation** and **Comedy**.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_1}
grid.arrange(category_likes_boxplot,
             top100_category_likes_boxplot,
             ncol=2)

summary(top100_trending$likes)
```
**Sports**, **Music** and **People & Blogs** pull quite ahead of other 
categories in terms of **views** with both some of the highest median and 
3<sup>rd</sup> quantile values.

Also, across most of the categories, the median number of views is higher in the 
Top 100 Trending Videos than those in the entire dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_2}
grid.arrange(category_dislikes_boxplot,
             top100_category_dislikes_boxplot,
             ncol=2)

summary(top100_trending$dislikes)
```
As we can see in the boxplot on the left for the entire datset, most of the 
categories have their median number of dislikes close to each other. But, in 
the Top 100 Trending Videos, we see a sharp decrease in it for the categories 
like **Education**, **Pets & Animals**, **News & Politics** etc. 

While for the popular categories like **Entertainment** and **Music** the median
 value of dislikes is one of the highest. Number of views and dislikes had a 
correlation of 0.71, so this is not very surprising.
 
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_3}
grid.arrange(category_commentCount_boxplot,
             top100_category_commentCount_boxplot,
             ncol=2)

summary(top100_trending$comment_count)
```
Again, **Sports** pulls ahead of other categories with **Gaming** having the 
second highest value in the Top 100 Trending Videos set. Also there is a sharp 
decrease in the category of **Education**.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_4}
grid.arrange(category_daysOnTrending_boxplot,
             top100_category_daysOnTrending_boxplot,
             ncol=2)

summary(top100_trending$days_on_trending)
```
The subsetting of the data is more readily visible for this feature that for any
 other feature with nearly all of the values for the categories being either 12,
 13 or 14.
 
Surprisingly, **Pets & Animals** is the category which trends for the longest 
time in the Top 100 Trending Videos. Whereas, the **Sports** category which was 
leading up until now in all the features has one of the lowest median value.


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=12, fig.height=6, Bivariate_Plots9_5}
grid.arrange(category_numberOfVideosTrending,
             top100_category_numberOfTrendingVideos_boxplot,
             ncol=2)

summary(top100_trending$number_of_videos_trending)
```
There is significant decrease in the number of trending videos of the **Sports**
 category between the entire dataset and the Top 100 Trending Videos.
 
While, the **Comedy** category maintains its domination even in the Top 100 
Trending Videos set with the highest median and 3<sup>rd</sup> quantile value 
closely followed by **News & Politics** and **Entertainment**.

___
___

<a id='bivariate-analysis'></a>

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

We have a high correlation between the **number of views** and **number of 
likes** both in the entire dataset (0.83) and in the Top 100 Trending Videos 
(0.85).

We also have a high correlation between the **number of dislikes** and **comment
count** in the entire dataset (0.83) but it drastically reduces in the Top 100 
Trending Videos (0.36).

The variation of the different features across categories varies quite widely 
depending upon the feature under investigation.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

In the Top 100 Trending Videos, the trend of **Entertainment** and **Music** 
having values across all the features still continues but is a lot less 
pronounced with other categories like **Film/Animation** and **People & Blogs**
coming close.
 
 **Pets & Animals** is surprising the category with the highest median, 
 1<sup>st</sup> and 3<sup>rd</sup> quantile values in the Top 100 Trending 
 Videos.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

In the Top 100 Trending Videos subset of the dataset, the **Pets & Animals** 
category has the highest median, 1<sup>st</sup> and 3<sup>rd</sup> quantile values.

### What was the strongest relationship you found?

The strongest correlation was between **views** and **likes** with 0.83 in the 
complete dataset and 0.85 in the Top 100 Trending Videos.

___
___

<a id='multivariate-plots-section'></a>

# Multivariate Plots Section

We have already seen the top trending channels. Here, we take a look at those 
channels and see how they are distributed across the different categories.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=number_of_videos_trending, fill=category_name)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("Number of trending videos")+
  scale_y_continuous(breaks=seq(0,140,20))

summary(
  yt_trending_by_number_of_videos_trending[1:25,]$number_of_videos_trending)
```

All of these channels post videos only of a specific category as we can see no 
channel having bars of different colours.

Furthermore, as we saw in the bivariate plots section, the most common 
categories are **Entertainment** and **Sports**.

**ESPN** having the highest number of trending videos (114). Both the mean and 
median of number of trending videos being 98.

In the bivariate plots section, we observed how the various features vary 
across different categories of videos. Now in this section, we can explore 
how these features vary across the Top 25 trending channels and see the 
classifications and conclusions of the bivariate plots section in action.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots_2}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=views, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("views")+
  scale_y_continuous(breaks=seq(0,100000000,1000000))

ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=views, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  scale_y_log10()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("log10(views)")

summary(yt_trending_by_number_of_videos_trending[1:25,]$views)
```

**ESPN**, **NBA**, **Netflix** and **Vox** are the top channels in terms of
number of videos on trending. But, **NFL**, **WIRED** and **WWE** are very ahead
 of these trending channels in terms of number of views.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots_3}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=likes, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("likes")+
  scale_y_continuous(breaks=seq(0,1000000,50000))

ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=likes, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  scale_y_log10()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("log10(likes)")

summary(yt_trending_by_number_of_videos_trending[1:25,]$likes)
```
Again, **ESPN**, **NBA**, **Netflix** and **Vox** are the top channels in terms 
of number of videos on trending. But, **First We Feast**, **NFL**, **The 
Tonight Show Starring Jimmy Fallon** and **WIRED**, and **WWE** are higher in 
terms of number of likes.


```{r echo=FALSE, warning=FALSE, Multivariate_Plots_4}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=dislikes, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("dislikes")+
  scale_y_continuous(breaks=seq(0,2000000,10000))

ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=dislikes, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  scale_y_log10()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("log10(dislikes)")

summary(yt_trending_by_number_of_videos_trending[1:25,]$dislikes)
```

Here we see, even **NFL** is coming close to the above four channels that we see 
topping the plots previously. We see a massive spike in the number of dislikes 
for **Washington Post** (132400) a **News & Politics** channel which is a order 
of magnitude higher than the median. Although, again, **ESPN** and **NFL** are 
still in the top 3.

(Considering the mean for comparision is not a good decision for this plot as 
there is a very large outlier to skew the mean)

```{r echo=FALSE, warning=FALSE, Multivariate_Plots_5}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=comment_count, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("comment count")

ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=comment_count, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  scale_y_log10()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("log10(comment count)")

summary(yt_trending_by_number_of_videos_trending[1:25,]$comment_count)
```

Again, we see large spikes for **CNN**, **WIRED** and **Washington Post** all of
 which are **News & Politics** channels. 

(The maximum value is again an order of magnitude higher than the median value) 

Hence, continuing the trend of people expressing their wide range of opinions 
not only through dislikes but also through comments.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots_6}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=days_on_trending, 
           fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("Days on trending")

summary(yt_trending_by_number_of_videos_trending[1:25,]$days_on_trending)
```

Although, the domination of **ESPN**, **NBA**, **Netflix** and **Vox** continues
 in terms of number of videos on trending, we see that only **NBA** channel's 
videos trend for a time longer than the average.

Also, channel's like **ABC News**, **Bon Appetit**, **Great Big Story** and 
**WIRED** all produce videos that trend for a longer time than the channel's 
mentioned above.

___
___

<a id='multivariate-analysis'></a>

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Looking at the top trending channels in terms of the number of videos trending, 
the same four channels **ESPN**, **NBA**, **Netflix** and **Vox** showed 
consistent domination across all the features except number of days on trending.

**News & Politics** channels in general, had very large values for features like 
**dislikes**, **comment count** and **days on trending** indicating strongs 
opinions of the people on the content posted by these channels.

### Were there any interesting or surprising interactions between features?

Features like **likes**, **views** etc. were generally higher for channels posting 
videos of the categories of **Entertainment** and **Music**.

Features like **dislikes**, **comment count** and **days on trending** were 
generally orders of magnitude higher for **News & Politics** channels.

___
___

<a id='final-plots-and-summary'></a>

# Final Plots and Summary

<a id='plot-one'></a>

### Plot One
```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Plot_One}
views_likes_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$views,
                               y=normalized_values$likes))+
  geom_point(alpha=0.2,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized likes")

dislikes_commentCount_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$dislikes,
                               y=normalized_values$comment_count))+
  geom_point(alpha=0.2,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.15)+
  ylim(0,0.15)+
  xlab("Normalized dislikes")+
  ylab("Normalized comment count")

grid.arrange(views_likes_plot,dislikes_commentCount_plot,ncol=2)
```

<a id='description-one'></a>

### Description One
We visualize the correlations between **views** and **likes** (both normalized; 
since numerically the number of views is always higher in the dataset) on the 
left and that between **dislikes** and **comment count** (also normalized) on 
the right.

One important trend we can observe from these correlations is that if a video 
has a high number of views then it also has a high number of likes indicating 
(not inferring) that the chance of being being popular due to bad publicity is 
low.

From the second plot, we can see that if a video has a high number of dislikes 
then it also tends to have a high comment count indicating that the video  may 
be controversial and hence the people are more likely to leave a comment.

Of course, these claims are just based on correlations and a hypothesis test is 
required to prove any causation.

<a id='plot-two'></a>

### Plot Two
```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=15, Plot_Two}
grid.arrange(category_views_boxplot,
             category_likes_boxplot,
             category_dislikes_boxplot,
             category_commentCount_boxplot,
             category_daysOnTrending_boxplot,
             category_numberOfVideosTrending,
             ncol = 2)

print("IQR of Shows for views:")
by(yt_trending$views,yt_trending$category_name=='Shows',IQR)

print("IQR of Shows for likes:")
by(yt_trending$likes,yt_trending$category_name=='Shows',IQR)

print("IQR of Shows for dislikes:")
by(yt_trending$dislikes,yt_trending$category_name=='Shows',IQR)

print("IQR of Shows for comment count:")
by(yt_trending$comment_count,yt_trending$category_name=='Shows',IQR)

print("IQR of Shows for days on trending:")
by(yt_trending$days_on_trending,yt_trending$category_name=='Shows',IQR)
```

<a id='description-two'></a>

### Description Two
Specifically, focussing on the **Shows** category of videos, we can see that of 
all the categories of videos across all of the features they have the least 
IQR value but also trend for the longest duration of time as can be seen from 
from the plots.

<a id='plot-three'></a>

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=dislikes, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("Number of dislikes")+
  scale_y_continuous(breaks=seq(0,2000000,10000))

ggplot(data=yt_trending_by_number_of_videos_trending[1:25,],
       aes(x=channel_title, y=comment_count, fill=number_of_videos_trending)) + 
    geom_bar(stat="identity",
             position=position_dodge(),
             size=.3)+
  scale_y_continuous(breaks=seq(0,50000,5000))+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Channel title")+
  ylab("Number of comments")
```

<a id='description-three'></a>

### Description Three
Between the two plots, we can see that videos posted by **Washington Post** both
 have the highest number of dislikes and one of the highest number of comments.

Looking at other channels in the above plots with notable values, **ESPN**, 
**NFL**, **CNN** and **WIRED** have some of the highest values for number of 
comments.

An interesting point to note is that, all of these channels post videos 
excludively of either the **News & Politics** or **Sports** category which are 
categories having supporters with different viewpoints and opinions and hence a 
high possibility of disagreement and this is clealy visible in the 
distribution of the data in the above plots.

___
___

<a id='reflection'></a>

# Reflection

The **YouTube Trending Statistics** is a daily record of the top trending videos
on the video sharing platform YouTube. The data used is that from the USA. It 
contains data about 23,362 (4,712 unique) videos across 13 features.
 
I began the analysis process by first preprocessing the data to remove some of 
the features and converting the dataset into a data.table type for quick 
access, merging etc. of the data. Some features like **days_on_trending** and 
**number_of_trending_videos** were derived from the data present in the dataset 
and added to it.

The analysis was carried out both over the entire dataset and over the Top 100 
Trending Videos (a subset of the entire dataset). In both cases, **views** and 
**likes** had a high correlation but **dislikes** and **comment count** had a 
high correlation only over the entire dataset but not in the Top 100 Trending 
Videos.

**Entertainment** and **Music** categories dominate the Top 100 Trending Videos
 in terms of the values of various features like **views**, **likes**,
**comment count** and **days on trending**. Also, videos of the category 
**Shows** trend for the longest time.

Further, looking at the top trending channels, positively indicative features 
like **views** and **likes** were high for the channels posting 
**Entertainment** and **Music** category of videos. While, features indicating 
controvery were high for channels posting videos of the **News & Politics** 
category.

This trend is easily observed online. Politics tends to be a category 
that generates a lot of controvery as people with different opinions 
come in contact. Whereas, music is in general a category that a majority 
of viewers have a similar opinion about.<br>
This facet of online interactions is quite accurately captured by this analysis.

I also tried to include a feature measuring the number of days a video took 
from its date of upload to start trending but ultimately could not due to not 
being able to preprocess the date format given in the dataset.

In addition to the analysis performed here, we could do a sentiment analysis 
using the title of the videos, video description and tags of the video. This can 
help in more accurately predicting the different features of the videos. This 
type of analysis can help understand how the perception of the audience
varies over different types of content and consequently help in the 
development of future content.