---
output: 
  html_document: 
    fig_height: 8
    fig_width: 12
    toc: yes
    toc_depth: 2
  pdf_document: 
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
editor_options: 
  chunk_output_type: inline
---
---
title: "EDA of Trending YouTube Video Statistics"
author: "Karra Anand"
date: "18 March 2018"
output: html_document
___
___

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(data.table)
library(corrplot)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
yt_trending = read.csv('USvideos.csv')
```

## About the dataset
YouTube (the world-famous video sharing website) maintains a list of the top 
trending videos on the platform. This dataset is a daily record of the top 
trending YouTube videos. <br>
This dataset includes several months (and counting) of data on daily trending 
YouTube videos. Data is included for the USA, with up to 200 listed trending 
videos per day. <br>
More details about this dataset are present in `About_datasat.txt` included with 
this project

___
___


## Preprocessing
Before we move on to plotting and analyzing the data, let us see if the data 
requires any cleaning.
<br>
<br>
The first 2 rows of each of the columns of the dataset are as follow:
```{r echo=FALSE,preproceessing}
head(yt_trending,n=2)
```
<br>

> In many of the following data cleaning steps only the code but not the output 
> is printed to prevent repeated printing of the same dataset with minor 
> modifications. The final dataset obtained after the data cleaning is printed 
> at the end of this section.

___

### Cleaning irrelevant columns
As we are only interested in exploratory analysis of the data, we remove the 
columns of **tags**, **thumbnail_link** and **description** since they are 
irrelevant to us.
```{r echo=FALSE, preprocessing_2}

yt_trending <- select(yt_trending,-tags,-thumbnail_link,
                      -description)

head(yt_trending)

dim(yt_trending)
```

___

### Converting into a data.table
From the dimensions of the dataframe in the above output, we have
the data for 23,362 videos(assuming there are no duplicates across 13 features; 
we shall investigate  this the following sections). Hence, we use the data.table
 data structure to store our data as it superior to a dataframe for add, remove,
 update, join etc. operations.
```{r echo=TRUE, preprocessing_3}
yt_trending <- data.table(yt_trending)
```
___

### Adding category names
Further we observe that, we have names of each of the **category_id** (as part
of `US_category_id.json` file). Hence, adding the column for **category_name**.

```{r echo=TRUE, preprocessing_4}
category_id <- c(1,2,10,15,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,
                  34,35,36,37,38,39,40,41,42,43,44)
category_name <- c("Film & Animation","Autos & Vehicles","Music",
                    "Pets & Animals","Sports","Short Movies",
                    "Travel & Events","Gaming","Videoblogging",
                    "People & Blogs","Comedy","Entertainment",
                    "News & Politics","Howto & Style","Education",
                    "Science & Technology",
                    "Nonprofits & Activism","Movies","Anime/Animation",
                    "Action/Adventure","Classics","Comedy",
                    "Documentary","Drama","Family","Foreign","Horror",
                    "Sci-Fi/Fantasy","Thriller","Shorts","Shows","Trailers")

category_id_names <- data.frame(category_id,category_name)

yt_trending <- merge(yt_trending,category_id_names)

yt_trending <- yt_trending[order(yt_trending$views,decreasing = TRUE),]
```

___

### Adding and sorting by days_on_trending
We are most interested here in analyzing the trends in the trending videos,
and we do not have a columnn to indicate the number of days a video was on 
trending, we add that column here.

```{r echo=TRUE, preprocessing_5}
days_on_trending_video_id <- (count(yt_trending,yt_trending$video_id))[1]
days_on_trending_num_of_days <- (count(yt_trending,yt_trending$video_id))[2]
days_on_trending <- data.table(days_on_trending_video_id,
                               days_on_trending_num_of_days)

setnames(days_on_trending,old="yt_trending$video_id",new="video_id")
setnames(days_on_trending,old="n",new="days_on_trending")

yt_trending <- merge(yt_trending,days_on_trending)

yt_trending <- yt_trending[order(yt_trending$days_on_trending,
                                 yt_trending$views,
                                 decreasing = TRUE),]
```

___

### Removing duplicate video entries
We sort the videos in decreasing order of their number of days on trending.

```{r echo=TRUE, preprocessing_6}
yt_trending <- unique(yt_trending, by = c("video_id"))

head(yt_trending)
```

___
___

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots_stats}
dim(yt_trending)
names(yt_trending)
summary(yt_trending)

likes_colour <- "chartreuse2"
dislikes_colour <- "red"
comment_count_colour <- "darkorchid"
days_on_trending_colour <- "deeppink"
views_colour <- "dodgerblue4"
```
As we can see from the above output, we have the data of 4,712 videos and their 
15 features.


```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots}
category_names_plot <- ggplot(yt_trending,aes(x=category_name))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Category")

category_names_plot

summary(yt_trending$category_name)
```
Plotting the number of videos as per their category name, we can see a large 
amount of variation in the number of videos in each category. We can clearly 
observe that **Entertainment** and **Music** are the top 2 categories with 1141 
and 585 videos respectively closely followed by **Howto & Style** with 436 
videos. We can also see their respective shares out of the total.
<br>
<br>
Also, some categories like **Action/Adventure**, **Anime/Animation** etc. have no 
videos.

___

```{r echo=FALSE, fig_width = 12, fig_height = 12, Univariate_Plots2, warning=FALSE}
likes_plot <- ggplot(yt_trending,aes(x=yt_trending$likes))+
  geom_histogram(bins=50,fill="chartreuse2")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of likes)")

dislikes_plot <- ggplot(yt_trending,aes(x=yt_trending$dislikes))+
  geom_histogram(bins=50,fill="red")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of dislikes)")

views_plot <- ggplot(yt_trending,aes(x=yt_trending$views))+
  geom_histogram(bins=50,fill="dodgerblue4")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of views)")

comments_plot <- ggplot(yt_trending,aes(x=yt_trending$comment_count))+
  geom_histogram(bins=50,fill="darkorchid")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of comments)")

grid.arrange(views_plot,likes_plot,dislikes_plot,comments_plot,ncol=1)

print("Summary of views feature:")
summary(yt_trending$views)

print("Summary of likes feature:")
summary(yt_trending$likes)

print("Summary of dislikes feature:")
summary(yt_trending$dislikes)

print("Summary of comments feature:")
summary(yt_trending$comment_count)
```
This plot shows the log10 of the number of **views**, **likes**, **dislikes** and 
**comments** on the videos.
<br>
<br>
We can observe a distribution similar to that of a normal distribution. Further,
we can see that, the mean number of views is more that that of any other 
attribute.
<br>
<br>
The **mean**, **median** and **maximum** number of view are 1,278,000, 331,600 
and 149,400,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of likes are 39,720, 7,726 and 
3,094,000 repectively.
<br>
<br>
The **mean**, **median** and **maximum** number of dislikes are 2,598, 302 and 
1,674,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of comments are 4,976, 888 and 
1,362,000 respectively.

___

```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots3}
comments_disabled_overall_plot <- ggplot(yt_trending,
                                         aes(x=yt_trending$comments_disabled))+
  geom_bar()+
  xlab("Comments disabled")

ratings_disabled_overall_plot <- ggplot(yt_trending,
                                        aes(x=yt_trending$ratings_disabled))+
  geom_bar()+
  xlab("Ratings disabled")

grid.arrange(comments_disabled_overall_plot,ratings_disabled_overall_plot,
             ncol=2)

print("Summary of videos with comments disabled")
summary(yt_trending$comments_disabled)

print("Summary of videos with ratings disabled")
summary(yt_trending$ratings_disabled)
```

So we have only 79 videos which have their comments disabled and only 26 videos 
with their ratings disabled.

___

```{r echo=FALSE, warning=FALSE, Univariate_Plots4}
ggplot(yt_trending,aes(x=yt_trending$days_on_trending))+
  geom_bar(fill="deeppink")+
  xlab("Days on trending")

summary(yt_trending$days_on_trending)
```
We can observe a bimodal type of distribution for the **days on trending**.
<br>
<br>
With videos trending for an average of about 5 days and a maximum of 14 days.

___

### Top 100 Trending Videos

```{r echo=FALSE, fig.width=12, fig.height=8, Univariate_Plots5,warning=FALSE}
top100_trending <- yt_trending[1:100]

top100_category_names_plot <- ggplot(top100_trending,aes(x=category_name))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Categories in Top 100 trending videos")

grid.arrange(top100_category_names_plot,category_names_plot)
```
Videos of some categories like **Auto & Vehicles**, **Comedy** etc. are not even
present in the top 100 trending videos.

<br>
```{r echo=FALSE, warning=FALSE, Univariate_Plots6}
top100_likes_plot <- ggplot(top100_trending,aes(x=top100_trending$likes))+
  geom_histogram(bins=50,fill="chartreuse2")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10 of number of likes on videos")

top100_dislikes_plot <- ggplot(top100_trending,aes(x=top100_trending$dislikes))+
  geom_histogram(bins=50,fill="red")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of dislikes)")

top100_views_plot <- ggplot(top100_trending,aes(x=top100_trending$views))+
  geom_histogram(bins=50,fill="dodgerblue4")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of views)")

top100_comments_plot <- ggplot(top100_trending,
                               aes(x=top100_trending$comment_count))+
  geom_histogram(bins=50,fill="darkorchid")+
  scale_x_log10(limits=c(1,100000000))+
  xlab("log10(number of comments)")


grid.arrange(top100_views_plot,
             top100_likes_plot,
             top100_dislikes_plot,
             top100_comments_plot,
             ncol=1)

print("Summary of views feature:")
summary(top100_trending$views)

print("Summary of likes feature:")
summary(top100_trending$likes)

print("Summary of dislikes feature:")
summary(top100_trending$dislikes)

print("Summary of comments feature:")
summary(top100_trending$comment_count)
```

For the top 100 trending videos we cannot see a definitive distribution as we 
could for these same features for the entire distribution.
<br>
<br>
The **mean**, **median** and **maximum** number of view are 3,411,000, 1,665,000 
and 45,940,000 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of likes are 105,600, 49,440 and 
822,000 repectively.
<br>
<br>
The **mean**, **median** and **maximum** number of dislikes are 5,052, 1,600 and 
165,100 respectively.
<br>
<br>
The **mean**, **median** and **maximum** number of comments are 11,980, 4,667 
and 203,900 respectively.
<br>
<br>
Comparing these statistics to those for the entire distribution, **all of these 
statistics are higher than those for the entire distribution**.

___

```{r echo=FALSE, warning=FALSE, Univariate_Plots7}
ggplot(top100_trending,aes(x=top100_trending$days_on_trending))+
  geom_bar(fill="deeppink")+
  xlab("Days on trending")

summary(top100_trending$days_on_trending)
```

Videos in the Top 100 trending videos trend for an average of 13 days with a 
maximum of 14 days.

___
___

## Univariate Analysis

### What is the structure of your dataset?
The dataset after preprocessing to remove duplicates, contains the data for 
4712 videos and 15 features (video_id, category_id, trending_date, title, 
channel_title, publish_time, views, likes, dislikes, comment_count, 
comments_disabled, ratings_disabled, video_error_or_removed, category_name, 
days_on_trending)
<br>
<br>
**Unordered factors: **category_id, category_name.
<br>
  **category_id: **1,2,10,15,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,
                  34,35,36,37,38,39,40,41,42,43,44
  <br>
  **category_name: **"Film & Animation","Autos & Vehicles","Music",
                    "Pets & Animals","Sports","Short Movies",
                    "Travel & Events","Gaming","Videoblogging",
                    "People & Blogs","Comedy","Entertainment",
                    "News & Politics","Howto & Style","Education",
                    "Science & Technology",
                    "Nonprofits & Activism","Movies","Anime/Animation",
                    "Action/Adventure","Classics","Comedy",
                    "Documentary","Drama","Family","Foreign","Horror",
                    "Sci-Fi/Fantasy","Thriller","Shorts","Shows","Trailers"
<br>
<br>
**Other observations:**

* **Entertainment** is the category with the highest number of video in 
trending with a total of 1141 videos.
* Average number of views on a trending video are 1,278,000.
* The maximum number of days for which a video stayed on trending is 14.
* Average number of views on a video in the top 100 trending videos is 
3,411,000.
* Most of the videos in the top 100 trending videos stay on trending for 13 
days.

___

### What is/are the main feature(s) of interest in your dataset?

The **number of views (views)**, **number of days the video is trending 
(days_on_trending)** and **category of the video (category_name)** are the main 
features of interest here.

___

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Although subject to the viewers' biases, number of likes, dislikes and comments 
can also help in understanding the video's position on the list of trending 
videos.

___

### Did you create any new variables from existing variables in the dataset?

Yes, the variable **category_name** was created to better interpret the 
category_id feature which has a direct correspondance with category_name. <br>
Also, the **days_on_trending** variable was created to keep track of the number 
of days the video was on trending.

___

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

One of the unusual observations were that there were no videos from some of the 
categories like **Action/Adventure**, **Anime/Animation** etc.

Some features like **tags**, **thubnail_link** and **description** were removed 
from the dataset as they took a lot of space on printing and were irrelevant to 
our analysis.

Log10 transform was applied in multiple plots to convey the scale and variation 
in the data as required.

Also, the same video which as trending on nultiple days was reduced to a single 
entry on the day it had the highest views.

___
___

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
corrplot_overall <- corrplot.mixed(corr = cor(yt_trending[,c("views",
                                         "likes",
                                         "dislikes",
                                         "comment_count",
                                         "days_on_trending"),with=F]))

#ggpairs(yt_trending[,c("views",
                       #"likes",
                       #"dislikes",
                       #"comment_count",
                       #"days_on_trending")])

```

The above correlation matrix helps in identifying some of the interesting trends
 in the data.
 
We have a high correlation between

* the **number of views** and **number of likes**, and 
* the **number of dislikes** and **comment count**.

<br>
But, before we plot scatter plots to visualize these correlations, we have to 
normalize the data ranges of the above mentioned four features.

<br>
After normalizing to be in the range of [0,1]. We get the following output:


```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots2}
range01 <- function(x){(x-min(x))/(max(x)-min(x))}

normalized_values <- select(yt_trending,views,likes,dislikes,comment_count)

normalized_values$views <- range01(normalized_values$views)
normalized_values$likes <- range01(normalized_values$likes)
normalized_values$dislikes <- range01(normalized_values$dislikes)
normalized_values$comment_count <- range01(normalized_values$comment_count)
head(normalized_values)

views_likes_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$views,
                               y=normalized_values$likes))+
  geom_point(alpha=0.2,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlab("Normalized views")+
  ylab("Normalized likes")

dislikes_commentCount_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$dislikes,
                               y=normalized_values$comment_count))+
  geom_point(alpha=0.2,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  xlab("Normalized dislikes")+
  ylab("Normalized comment count")

grid.arrange(views_likes_plot,dislikes_commentCount_plot,ncol=2)
```

Using suitable limits for the X and Y axis:

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots3}
views_likes_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$views,
                               y=normalized_values$likes))+
  geom_point(alpha=0.2,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized likes")

dislikes_commentCount_plot <- ggplot(normalized_values,
                           aes(x=normalized_values$dislikes,
                               y=normalized_values$comment_count))+
  geom_point(alpha=0.2,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.15)+
  ylim(0,0.15)+
  xlab("Normalized dislikes")+
  ylab("Normalized comment count")

grid.arrange(views_likes_plot,dislikes_commentCount_plot,ncol=2)
```

We can clearly observe the correlation that we found out previously using 
the correlation matrix.


We can see the variation in the features of **likes**, **dislikes**, **comment 
count** and **days on trending** in the following plots.

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=8, Bivariate_Plots4}
likes_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$likes,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(likes)")

dislikes_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$dislikes,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="red")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(dislikes)")

comments_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$comment_count,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(comment count)")

daysOnTrending_views_plot <- ggplot(yt_trending,
       aes(y=yt_trending$days_on_trending,x=yt_trending$views))+
  geom_point(alpha=0.05,colour="deeppink")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  xlab("log10(views)")+
  ylab("days on trending")

grid.arrange(likes_views_plot,
             dislikes_views_plot,
             comments_views_plot,
             daysOnTrending_views_plot,
             ncol=2)
```

The notable but not considerable correlations like in between **views** and 
**dislikes** and that between **views** and **comment count** are visible here.

As the number of views increases, the other features also increase which is in 
agreement with our calculated statistics in the univariate plots section.

Moreover, only a small number of videos trend for 10 days or more.
<br>


```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=15, Bivariate_Plots5}
category_views_boxplot <- ggplot(yt_trending,aes(x=category_name,y=views))+
  geom_boxplot(alpha=0.1,colour="dodgerblue4")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(views)")+
  xlab("category name")

category_likes_boxplot <- ggplot(yt_trending,aes(x=category_name,y=likes))+
  geom_boxplot(alpha=0.1,colour="chartreuse2")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(likes)")+
  xlab("category name")

category_dislikes_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                    y=dislikes))+
  geom_boxplot(alpha=0.1,colour="red")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(dislikes)")+
  xlab("category name")

category_commentCount_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                        y=comment_count))+
  geom_boxplot(alpha=0.1,colour="darkorchid")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(comment count)")+
  xlab("category name")

category_daysOnTrending_boxplot <- ggplot(yt_trending,aes(x=category_name,
                                                          y=days_on_trending))+
  geom_boxplot(alpha=0.1,colour="deeppink")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  ylab("days on trending")+
  xlab("category name")

grid.arrange(category_views_boxplot,
             category_likes_boxplot,
             category_dislikes_boxplot,
             category_commentCount_boxplot,
             category_daysOnTrending_boxplot,
             ncol = 2)
```

The categories of **Entertainment** and **Music** have very high values for all 
the features across the board.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

___

### Top 100 Trending Videos

In the univariated plots and analysis sections, we observed that the trends in 
the entire dataset were more pronounced in the Top 100 Trending Videos.

We check if this is true for the correlations that we observed at the beginning 
of the bivariate plots section.

```{r echo=FALSE, Bivariate_Plots6}
corrplot_top100 <- corrplot.mixed(corr = cor(top100_trending[,c("views",
                                         "likes",
                                         "dislikes",
                                         "comment_count",
                                         "days_on_trending"),with=F]))

# ggpairs(top100_trending[,c("views",
#                        "likes",
#                        "dislikes",
#                        "comment_count",
#                        "days_on_trending")])
```


The changes in the correlations from those of the entire dataset are:

* Minor increase from 0.81 to 0.85 between **views** and **likes** 
* Drastic reduction from 0.83 to 0.36 between **dislikes** and **comment count**
* Of the other correlations, some has increased and some have decreased

Again before plotting, we first normalize the values of the features to be in 
the range [0,1].

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots7}
normalized_top100_trending <- select(top100_trending,
                                     views,
                                     likes,
                                     dislikes,
                                     comment_count)

normalized_top100_trending$views <- range01(top100_trending$views)
normalized_top100_trending$likes <- range01(top100_trending$likes)
normalized_top100_trending$dislikes <- range01(top100_trending$dislikes)
normalized_top100_trending$comment_count <- 
  range01(top100_trending$comment_count)

head(normalized_top100_trending)


top100_views_likes_plot <- ggplot(normalized_top100_trending,
                           aes(x=normalized_top100_trending$views,
                               y=normalized_top100_trending$likes))+
  geom_point(alpha=0.5,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized likes")

top100_views_dislikes_plot <- ggplot(normalized_top100_trending,
                           aes(x=normalized_top100_trending$views,
                               y=normalized_top100_trending$dislikes))+
  geom_point(alpha=0.5,colour="red")+
  geom_abline(slope=1,intercept=0)+
  xlim(0,0.3)+
  ylim(0,0.3)+
  xlab("Normalized views")+
  ylab("Normalized dislikes")

grid.arrange(top100_views_likes_plot,top100_views_dislikes_plot,ncol=2)
```

We can observe the correlation that we found from the correlation matrix above.

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=8, Bivariate_Plots8}
top100_likes_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$likes,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="chartreuse2")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(likes)")

top100_dislikes_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$dislikes,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="red")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(dislikes)")

top100_comments_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$comment_count,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="darkorchid")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("log10(views)")+
  ylab("log10(comment count)")

top100_daysOnTrending_views_plot <- ggplot(top100_trending,
       aes(y=top100_trending$days_on_trending,x=top100_trending$views))+
  geom_point(alpha=0.5,colour="deeppink")+
  geom_abline(slope=1,intercept=0)+
  scale_x_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  xlab("log10(views)")+
  ylab("days on trending")

grid.arrange(top100_likes_views_plot,
             top100_dislikes_views_plot,
             top100_comments_views_plot,
             top100_daysOnTrending_views_plot,
             ncol=2)
```

The data points are further away from the y=x line and hence overall, with 
**views** the other features are less correlated.

Also, videos in the Top 100 Trending Videos trend for either 12, 13 or 14 days 
only.

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=25, Bivariate_Plots9}
top100_category_views_boxplot <- ggplot(top100_trending,aes(x=category_name,y=views))+
  geom_boxplot(alpha=0.1,colour="dodgerblue4")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(views)")+
  xlab("category name")


top100_category_likes_boxplot <- ggplot(top100_trending,aes(x=category_name,y=likes))+
  geom_boxplot(alpha=0.1,colour="chartreuse2")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(likes)")+
  xlab("category name")

top100_category_dislikes_boxplot <- ggplot(top100_trending,aes(x=category_name,
                                                    y=dislikes))+
  geom_boxplot(alpha=0.1,colour="red")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(dislikes)")+
  xlab("category name")

top100_category_commentCount_boxplot <- ggplot(top100_trending,aes(x=category_name,
                                                        y=comment_count))+
  geom_boxplot(alpha=0.1,colour="darkorchid")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  ylab("log10(comment count)")+
  xlab("category name")

top100_category_daysOnTrending_boxplot <- ggplot(top100_trending,aes(x=category_name,
                                                          y=days_on_trending))+
  geom_boxplot(alpha=0.1,colour="deeppink")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  scale_y_log10()+
  scale_y_continuous(breaks=seq(0,14,1))+
  ylab("days on trending")+
  xlab("category name")

grid.arrange(top100_category_views_boxplot,
             category_views_boxplot,
             top100_category_likes_boxplot,
             category_likes_boxplot,
             top100_category_dislikes_boxplot,
             category_dislikes_boxplot,
             top100_category_commentCount_boxplot,
             category_commentCount_boxplot,
             top100_category_daysOnTrending_boxplot,
             category_daysOnTrending_boxplot,
             ncol = 2)
```

The plots from **Top 100 Trending Videos are on the left** and those from the 
**entire dataset are on the right**.

The subsetting of data for the Top 100 Trending Videos is clearly visible with 
the absence of a lot of data points.

The trend of **Entertainment** and **Music** having values across all the 
features still continues but is a lot less pronounced with other categories like
 **Film/Animation** and **People & Blogs** coming close.
 
 **Pets & Animals** is surprising the category with the highest median, 
 1<sup>st</sup> and 3<sup>rd</sup> quantile values in the Top 100 Trending 
 Videos.

___
___

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

We have a high correlation between:

* the **number of views** and **number of likes**, and 
* the **number of dislikes** and **comment count**

in the dataset containing the all of the data but not in the Top 100 Trending 
Videos.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

**Nonprofits & Activism** have the lowest median and 1<sup>st</sup> 
quantile values for nearly all the features.

**Shows** show a very small difference between the 3<sup>rd</sup> and 
1<sup>st</sup> quantile for all the features. They also have the highest median,
1<sup>st</sup> and 3<sup>rd</sup> quantile value for **days on trending**.

In the Top 100 Trending Videos, the trend of **Entertainment** and **Music** 
having values across all the features still continues but is a lot less 
pronounced with other categories like **Film/Animation** and **People & Blogs**
coming close.
 
 **Pets & Animals** is surprising the category with the highest median, 
 1<sup>st</sup> and 3<sup>rd</sup> quantile values in the Top 100 Trending 
 Videos.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

In the Top 100 Trending Videos subset of the dataset, the **Pets & Animals** 
category has the highest median, 1<sup>st</sup> and 3<sup>rd</sup> quantile values.

### What was the strongest relationship you found?

The strongest correlation was between **views** and **likes** with 0.83 in the 
complete dataset and 0.85 in the Top 100 Trending Videos.

___
___
